Meta-analytic approaches and 'new' effect sizes to account for heterogeneous effects in comparative physiology

### 2021-07-08 MS/ms.Rmd
Small changes to opening sentences

The goal of our review is to briefly overview the process of meta-analysis, making a case for why heterogenity in a meta-analysis is a good thing for answering fundamental reserch problems and guiding research directions. We introduce the idea of 'nuisance heterogenity', showing comparative physiologists that it can be comfortably dealt with be re-formalising many existing effect sizes measures and/or by using multilevel meta-regression models. Indeed, a combination of these approaches might often be ideal. We first conduct a survey of published meta-analyses in the field of comparative physiology to gauge methodological and meta-analytic practices -- assessing what types of effect sizes are commonly used, how often effect size estimates are impacted by continuous treatment differences and whether meta-analytic models (i.e., models that account for sampling variance) are commonly applied. We then formalise alternative effect sizes and their associated sampling variance to provide comparative physiologists opportunities to explicitly incorporate treatment differences into the effect size estimates to ease their interpretation. We then describe how more complex treatment differences, such as dosage differences (which may have non-linear effects) can be accommodated using multilevel meta-regression models. We hope to show how expanding the meta-analytic toolkit for comparative physiologists will provide new opportunities to address global scale problems that attempt to understand how organisms will cope with rapidly changing environments in the future.

@@

where $CV$ is the coefficient of variation defined as $SD / R$. Whether one choose to use SD or CV-based effect size depends on questions at hand [discussed in @Nakagawa2015 and @Senior2020].

@@

$$
\sigma_{m}^2 = \sum w_{j}\left( k-1\right) / \left[ \left( \sum w_{j}\right)^2  + \sum w_{j}^2\right]
$$

where *k* is the number of studies and the weights, $w_{j} = \frac{1}{v_{j}}$, can be calcualted as the inverse of the sampling variance ($v_{j}$) for each effect size, *j*.

>>>>

The goal of our review is to briefly overview the process of meta-analysis, making a case for why heterogenity in a meta-analysis is a good thing for answering fundamental reserch problems and guiding research directions. We introduce the idea of 'nuisance heterogenity', showing comparative physiologists that it can be comfortably dealt with by re-formalising many existing effect sizes measures and/or by using multilevel meta-regression models. Indeed, a combination of these approaches might often be ideal. We first conduct a survey of published meta-analyses in the field of comparative physiology to gauge methodological and meta-analytic practices -- determining what types of effect sizes are commonly used, how often effect size estimates are impacted by 'nuisance heterogeneity' and whether meta-analytic models (i.e., models that account for sampling variance) are commonly applied. We then formalise alternative effect sizes and their associated sampling variance to provide comparative physiologists opportunities to explicitly incorporate treatment differences into the effect size estimates to ease their interpretation. We then describe how more complex treatment differences, such as dosage differences (which may have non-linear effects) can be accommodated using multilevel meta-regression models. We hope to show how expanding the meta-analytic toolkit for comparative physiologists will provide new opportunities to address global scale problems that attempt to understand how organisms will cope with rapidly changing environments in the future.

@@

where $CV$ is the coefficient of variation defined as $SD / R$. Whether one choose to use SD or CV-based effect size depends on questions at hand [See @Nakagawa2015; @Senior2020].

@@

$$
\sigma_{m}^2 = \sum w_{i}\left( k-1\right) / \left[ \left( \sum w_{i}\right)^2  + \sum w_{i}^2\right]
$$

where *k* is the number of studies and the weights, $w_{i} = \frac{1}{v_{i}}$, can be calcualted as the inverse of the sampling variance ($v_{i}$) for each effect size, *i*.
